Statically Typed String Sanitation Inside a Python

### 2014-08-19 sanitation-psp14/sanitation-psp14-short.tex
Fixed layout of the paper.

@@

\section{Regular String Types, Minimally}\label{calculus}
\renewcommand{\grammarlabel}[2]{#1\hfill#2}

\newcommand{\sctx}{\Psi} % Context for external typing
\newcommand{\tctx}{\Theta} % Context for internal typing

\newcommand{\ereduces}{\Downarrow}

\begin{figure}[t]
\begin{grammar}
<$r$> ::= $\epsilon$ | $.$ | $a$ | $r \cdot r$ | $r + r$ | $r*$ \hfill $a \in \Sigma$

\begin{figure}[t]
\begin{grammar}

<$e$> ::= 
      $x$ \hfill source terms \alt 
      $\lambda x . e$ \alt
      $e(e)$ \alt
      $\strin{s}$ \hfill $s \in \Sigma^{*}$ \alt
      $\rsconcat{e}{e}$ \alt
      $\sisubst{r}{e}{e}$ \alt
      $\rcoerce{r}{e}$ \alt
      $\rcheck{r}{e}{x.e}{e}$
\caption{Syntax of $\lambda_{RS}$.}\label{fig:lamrs}
\end{grammar}
\end{figure}

@@

@@

This section ultimately establishes that the security theorem for $\lambdas$ is preserved under compilation.

\begin{figure}\label{fig:lambdas}
\small
$\fbox{\inferrule{}{\sctx \vdash e : \sigma}}$
~~~~$\sctx ::= \emptyset \pipe \sctx, x : \sigma$
\begin{mathpar}
\inferrule[S-T-Var]
{ x:\sigma \in \sctx }
{ \sctx \vdash x:\sigma}

\caption{Typing rules for $\lambdas$. 
The typing context $\sctx$ is standard.}
\label{fig:etyping}
\end{figure}

\end{mathpar}
\caption{Big step semantics for $\lambdas$}.
\label{fig:eeval}
\end{figure}

%
% Begin typing/translation relation
%

% Begin Grammar 
% Removes the < > from grammar productions.
\renewcommand{\grammarlabel}[2]{#1\hfill#2}

% \begin{figure}[t]
% \begin{grammar}
% <r> ::= $\epsilon$ | $.$ | $a$ | $r \cdot r$ | $r + r$ | $r*$ \hfill $a \in \Sigma$

% \caption{Regular expressions over the alphabet $\Sigma$.}
% \label{fig:regex}
% \end{grammar}
% \end{figure}

% \begin{figure}[h]
% \begin{grammar}

% <$\psi$> ::=	$\psi \rightarrow \psi$			\hfill	source types					\alt
% $\stringin{r}$

% <S> ::= 
%       $\lambda x . e$ \hfill source terms \alt
%       $e e$ \alt
%       $\strin{s}$ \hfill $s \in \Sigma^{*}$ \alt
%       $\rsconcat{S}{S}$ \alt
%       $\sisubst{r}{S}{S}$ \alt
%       $\coerce{r}{S}$
% \caption{Syntax for the string sanitation fragment of our source language, $\lambdas$.}
% \end{grammar}
% \end{figure}

\begin{figure}[t]\label{fig:lambdap}
\small
$\fbox{\inferrule{}{\tctx \vdash \iota : \tau}}$
~~~~$\tctx ::= \emptyset \pipe \tctx, x : \tau$

@@

\inferrule[P-T-Check]
{\tctx \vdash \iota_r : \regex \\ \tctx \vdash \iota_1 : \str \\ \tctx, x:\str \vdash \iota_2 : \sigma \\ \tctx \vdash \iota_3 : \sigma}
{\tctx \vdash \tcheck{\iota_r}{\iota_1}{x . \iota_2}{\iota_3} : \sigma}
\end{mathpar}
\caption{Typing rules for $\lambdap$.
The typing context $\tctx$ is standard.}
\label{fig:ityping}
\end{figure}

@@

\inferrule[P-E-Check-NotOK]
{\iota \treduces \tstr{s} \\ s \not \in \lang{r}}
{\tcheck{\rx{r}}{\iota}{x.\iota_1}{\iota_2} \treduces \iota_2}
\end{mathpar}
\caption{Big step semantics for of $\lambdap$}.
\label{fig:ieval}
\end{figure}

\newpage
\ 
\newpage
\ 
\newpage

\begin{figure}[h]\label{fig:trans}

@@

\end{mathpar}
\caption{Translation from source terms ($e$) to target terms ($\iota$).
The translation is type-directed.}
\end{figure}

@@

\subsection{An Example}\label{whatev}

@@

\label{fig:impl}
\begin{figure}
\begin{lstlisting}
class string_in(atlang.Type):
  def __init__(self, rx):
    rx = rx_normalize(rx)
    atlang.Type.__init__(idx=rx)

@@

  def check_Coerce(self, ctx, node, other_t):
    # coercions can only be defined between 
    # types with the same type constructor, 
    if rx_sublang(other_t.idx, self.idx):
      return other_t
    else: raise atlang.TypeError("...", node)
\end{lstlisting}
\vspace{-10px}
\caption{Implementation of the \texttt{string\_in} type constructor in atlang.}
\label{fig:atimpl}\end{figure}

>>>>

\newcommand{\sctx}{\Psi} % Context for external typing
\newcommand{\tctx}{\Theta} % Context for internal typing
\newcommand{\ereduces}{\Downarrow}

@@

\section{Regular String Types, Minimally}\label{calculus}

@@

In this section, we define $\lambdas$, describe each operation, and finally prove both type safety and a correctness result for the language.
The grammar for $\lambdas$ is defined in Figure \ref{fig:glambdas}. Typing rules \textsc{S-T-} are defined in Figure \ref{fig:slambdas} and a big-step semantics
is defined in Figure \ref{fig:dlambdas}
\footnote{For the reader's convienance, a single sheet containing figures \ref{fig:glambdas}-\ref{fig:tr} is available at http://nfulton.org/strings/printout.pdf \todo{post this}}.

\renewcommand{\grammarlabel}[2]{#1\hfill#2}
\begin{figure}[h]
\begin{grammar}
<$r$> ::= $\epsilon$ | $.$ | $a$ | $r \cdot r$ | $r + r$ | $r*$ \hfill $a \in \Sigma$

<$e$> ::= 
      $x$ \hfill source terms \alt 
      $\lambda x . e$ \alt
      $e(e)$ \alt
      $\strin{s}$ \hfill $s \in \Sigma^{*}$ \alt
      $\rsconcat{e}{e}$ \alt
      $\sisubst{r}{e}{e}$ \alt
      $\rcoerce{r}{e}$ \alt
      $\rcheck{r}{e}{x.e}{e}$
\caption{Syntax of $\lambda_{RS}$.}
\label{fig:glambdas}
\end{grammar}
\end{figure}

\begin{figure}[h]
\small
$\fbox{\inferrule{}{\sctx \vdash e : \sigma}}$
~~~~$\sctx ::= \emptyset \pipe \sctx, x : \sigma$
\begin{mathpar}
\inferrule[S-T-Var]
{ x:\sigma \in \sctx }
{ \sctx \vdash x:\sigma}

\caption{Typing rules for $\lambdas$. 
The typing context $\sctx$ is standard.}
\label{fig:slambdas}
\end{figure}

\end{mathpar}
\caption{Big step semantics for $\lambdas$}.
\label{fig:dlambdas}
\end{figure}

@@

The grammar of $\lambdap$ is defined in Figure~\ref{fig:lcsSyntax}.
The typing rules \textsc{P-T-} are defined in Figure~\ref{fig:slambdap}
and a big-step semantics is defined by the rules \textsc{P-E-} in Figure~\ref{fig:dlambdap}.

\renewcommand{\grammarlabel}[2]{#1\hfill#2}

\begin{figure}[h]\label{fig:lambdap}
\small
$\fbox{\inferrule{}{\tctx \vdash \iota : \tau}}$
~~~~$\tctx ::= \emptyset \pipe \tctx, x : \tau$

@@

\inferrule[P-T-Check]
{\tctx \vdash \iota_r : \regex \\ \tctx \vdash \iota_1 : \str \\ \tctx, x:\str \vdash \iota_2 : \sigma \\ \tctx \vdash \iota_3 : \sigma}
{\tctx \vdash \tcheck{\iota_r}{\iota_1}{x . \iota_2}{\iota_3} : \sigma}
\end{mathpar}
\caption{Typing rules for $\lambdap$.
The typing context $\tctx$ is standard.}
\label{fig:slambdap}
\end{figure}

@@

\inferrule[P-E-Check-NotOK]
{\iota \treduces \tstr{s} \\ s \not \in \lang{r}}
{\tcheck{\rx{r}}{\iota}{x.\iota_1}{\iota_2} \treduces \iota_2}
\end{mathpar}
\caption{Big step semantics for of $\lambdap$}.
\label{fig:dlambdap}
\end{figure}

The translation is defined by the rules \textsc{Tr-} in Figure~\ref{fig:tr}.
This section ultimately establishes that the security theorem for $\lambdas$ is preserved under compilation.

\begin{figure}[h]

@@

\end{mathpar}
\caption{Translation from source terms ($e$) to target terms ($\iota$).
The translation is type-directed.}
\label{fig:tr}
\end{figure}

\subsubsection{Preservation of Security}

@@

%\subsection{An Example}\label{whatev}

@@

\begin{figure}
\begin{lstlisting}
class string_in(atlang.Type):
  def __init__(self, rx):
    rx = rx_normalize(rx)
    atlang.Type.__init__(idx=rx)

@@

  def check_Coerce(self, ctx, node, other_t):
    # coercions can only be defined between 
    # types with the same type constructor, 
    if rx_sublang(other_t.idx, self.idx):
      return other_t
    else: raise atlang.TypeError("...", node)
\end{lstlisting}
\vspace{-10px}
\caption{Implementation of the \texttt{string\_in} type constructor in atlang.}
\label{fig:impl}

\label{fig:atimpl}\end{figure}
